---
title: "Aug-26 Demo"
output: word_document
---

```{r}
prof.data<-read.csv('C:/Projects/R/bizanalytics/data/Professor.csv')
```

## Perform Univariate Descriptive Analysis:
## 1. Summarize a categorical variable 

Let's summarize **Gender** variable.

### 1.1) Summarize by frequency table
```{r}
freq_table<-table(summary(prof.data$Gender))
freq_table
```

### 1.2) Generate a barplot
```{r}
barplot(freq_table,
        main="Summary of Gender in Professor Data",
        xlab="Gender")
```

### 1.3) Generate a Pie Chart
```{r}
pie(table(freq_table))
```


## 2. Summarize a numeric variable

Let's summarize **Salary** variable.

### 2.1) Summarize by numbers
We can use summary() to get a quick preview:
```{r}
summary(prof.data$Salary)
```
We can also apply different functions to get individual measure:
mean():
```{r}
mean(prof.data$Salary)
```

median():
```{r}
median(prof.data$Salary)
```

min() and max()
```{r}
min(prof.data$Salary)
max(prof.data$Salary)
```

1st quartile and 3rd quartile
```{r}
quantile(prof.data$Salary,0.25)
quantile(prof.data$Salary,0.75)
```

standard deviation and variance
```{r}
sd(prof.data$Salary)
var(prof.data$Salary)
```

### Summarize by chart (Histogram)
```{r}
with(prof.data,
     hist(Salary,breaks=15)
     )
```

### Summarize by chart (Boxplot)
```{r}
with(prof.data,
     boxplot(Salary))
```

