---
title: "practice"
author: "Justin Markey"
date: "8/26/2021"
---
```{r}
data1 = read.csv("Professor.csv")

x = (data1$Exper)
hist(x)
```

scatterplot
```{r}

scatterplot(data=data1,
            x=Exper,
            y=Salary)

```

Show box plot
```{r}
with(data1,
     plot(x=Salary, y=Gender))
```

conditional ##scatterplot, use car package

```{r}
library(car)

```


rank the data and make a boxplot
```{r}
library(dplyr)

data1$Exper_Rank = with(data1, ntile(Exper,4))

##plot it

with(data1,
     plot(Salary~Exper_Rank))

```

find target variable to be categorical
mosiacplot
make a new cateogorical var called High_Salary if salary is greater than 90 otherwise it is 0
```{r}
data1$High_Salary = with(data1,
                         ifelse(Salary>90,1,0))
```

Produce contingency table and store it as a var
```{r}

contingency = with(data1, table(Gender, High_Salary))
```

Graph mosaic
```{r}
mosaicplot(contingency)
```


Y: categorical, x: numeric

change x into categorical data{ranks}, so that its categorical vs categorical
produce table to make a mosiacplot

```{r}
contingency2 =with(data1,table(Exper_Rank , High_Salary))
mosaicplot(contingency2)
```

