---
title: "Class Demo 20210916"
output: word_document
---

# Illustration of Regression with CEO Compensation data
```{r}
ceo.data<-read.csv("Executive_Compensation.csv")
```

## Build separate simple linear regression
```{r}
model.1<-lm(data=ceo.data,Compensation~Adj.ROA)
model.2<-lm(data=ceo.data,Compensation~Adj.Returns)
model.3<-lm(data=ceo.data,Compensation~Total.Assets)

summary(model.1)
summary(model.2)
summary(model.3)
```

## Build a multiple regression
```{r}
model.4<-lm(data=ceo.data,
            Compensation~.)
summary(model.4)
```

## Predictions with linear regression
## Type 1: in-sample prediction
```{r}
predict(model.1)
```

## Type 2: out-of-sample prediction
Let's assume we predict for three ceos with the following information
CEO1: Adj.ROA=0.5, Adj.Returns=0.5, Total.Assets=16
CEO2: Adj.ROA=-0.2, Adj.Returns=-0.3, Total.Assets=400
CEO3: Adj.ROA=0.2, Adj.Returns=0.3, Total.Assets=50

```{r}
test.data<-data.frame(
  Adj.ROA=c(0.5,-0.2,0.2),
  Adj.Returns=c(0.5,-0.3,0.3),
  Total.Assets=c(16,400,50)
)
predict(model.1,newdata=test.data)
predict(model.4,newdata=test.data)
```
